```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## LIBRA Example 

Compiled: June 15, 2021.

PPJI metric: Clustering preservation computation between LIBRA integrated subspace clustering and original omics RNA/ATAC clustering.


```{r prepare the env}
#options(warn=-1)

###########################
#Prepare enviroment
set.seed(1234567)
options(stringsAsFactors = FALSE)

#Commented resourcess will be required for training the model
#library("keras")
#K <- keras::backend()
#tensorflow::tf$compat$v1$disable_eager_execution()
#print("backend mounted")

#Extra resources for manage data, clustering and visualization
library("Seurat")
library("scclusteval")
library("ggplot2")

print("Requirements loaded")
gc()
```

Setting the workspace. Loading the output of the pre-trained model as well as pre-processed RNA and ATAC Seurat objects.

```{r set working dir}
setwd("/your_path")
load("LIBRA_output.RData")
print("Loaded LIBRA output")
```

Pairwise Jaccard distance between reference omic RNA (y-axis) clusters and integrated cluster obtain in LIBRA (x-axis)

```{r PJI}
jacard=PairWiseJaccardSetsHeatmap(rna_preprocessed@active.ident, LIBRA@active.ident,
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
jacard
```

PPJI summarization for corresponding dataset

```{r PPJI}
ppji_score = c()
jacard_distance_matrix = jacard@matrix
is.na(jacard_distance_matrix) = jacard_distance_matrix==0
jacard_distance_matrix = jacard_distance_matrix[,colSums(is.na(jacard_distance_matrix))<nrow(jacard_distance_matrix)]
ppji_score = append(ppji_score, mean(apply(jacard_distance_matrix,1,sum, na.rm=TRUE)))
print(ppji_score)
```



